Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-10-31 concentration_argument.tex
fully merged compations of Nikolaos plus some small edits on section 2

In particular, the average degree of a node at height $y$ is approximately Poisson with mean $\xi_{\alpha,\nu} e^{y/2}$ while we are interested in those nodes with degree $k_n$. Since Poisson random variables are well concentrated around their mean, for our analysis we would therefore like to be able the integration to intervals where $e^{y/2} \approx k_n$. Moreover, we want to be able to do this for all three Poisson models $G_{\HP,n}$, $G_{\Pcal,n}$ and $G_\Pcal$ as well as simultaneously deal with the case where $k_n \to \infty$ and $k_n = \bigT{1}$. In this section we will establish such a result. We start with a concentration lemma for the infinite model (Lemma~\ref{lem:concentration_argument}) and explain in Remark~\ref{rmk:concentration_argument} how such a result will be used throughout the paper. To obtain similar results for the other two models we have to first analyze the average degree in these models which we do in Sections~\ref{ssec:average_degree_HP_n} and~\ref{ssec:average_degree_P_n}. We conclude this section with a general result that allows us to extend the concentration lemma for the infinite model to the hyperbolic random graph and finite box model.

>>>>

In particular, the average degree of a node at height $y$ is approximately Poisson with mean $\xi_{\alpha,\nu} e^{y/2}$ and we are mostly interested in those nodes with degree $k_n$. Since Poisson random variables are well concentrated around their mean, for our analysis we would therefore like to be able the restrict integration to intervals where $e^{y/2} \approx k_n$. Moreover, we want to be able to do this for all three Poisson models $G_{\HP,n}$, $G_{\Pcal,n}$ and $G_\Pcal$ as well as simultaneously deal with the case where $k_n \to \infty$ and $k_n = \bigT{1}$. In this section we will establish such a result. We start with a concentration lemma for the infinite model (Lemma~\ref{lem:concentration_argument}) and explain in Remark~\ref{rmk:concentration_argument} how such a result will be used throughout the paper. To obtain similar results for the other two models we have to first analyze the average degree in these models which we do in Sections~\ref{ssec:average_degree_HP_n} and~\ref{ssec:average_degree_P_n}. We conclude this section with a general result that allows us to extend the concentration lemma for the infinite model to the hyperbolic random graph and finite box model.

### 2018-10-31 introduction.tex
fully merged compations of Nikolaos plus some small edits on section 2

In the next two subsections \ref{ssec:hyperbolic_model} and \ref{ssec:local_clustering} of the introduction, we give the definitions of the hyperbolic random graph and of the local clustering coefficient and function as used in this paper. After this we present our main results in Section \ref{ssec:main_results} and discuss several insights that can be obtained from them and in Section \ref{ssec:simulations} show simulations which confirm these results. Section \ref{sec:proof_outline} contains a detailed high level outline of the proofs of our main results, split up into different propositions. The proofs of these propositions can be found in the last four sections. Section~\ref{sec:concentration_argument} contains a crucial technical result that will be used in the other sections.

@@

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
All our results for clustering in the hyperbolic random graph are valid for any sequence $k_n$ such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the optimal scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_local_clustering_general} we have that $c_{\H,n}(k_n) = 0$ if $N_{\H,n}(k_n) = 0$. Hence, it follows by Markov's inequality that for any positive function $f$
\begin{align*}
	\Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	&\ge \Prob{N_{\H,n}(k_n) = 0} \ge 1 - \Exp{N_{\H,n}(k_n)}.
\end{align*}
We shall later establish (see Lemma [??] \PvdH{Find the reference for this Lemma}) that $\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}$. Therefore if $k_n$ is such that $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$ we have
\[
	\lim_{n \to \infty} n k_n^{-(2\alpha + 1)} 
	= \lim_{n \to \infty} \left(n^{-\frac{1}{2\alpha + 1}} k_n\right)^{-(2\alpha + 1)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	\ge \lim_{n \to \infty} 1 - \Exp{N_{\H,n}(k_n)} = \lim_{n \to \infty} 1 - \bigT{n k_n^{-(2\alpha + 1)}} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:clustering_coefficient_hyperbolic} to hold as soon as $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$.

>>>>

In the next two subsections \ref{ssec:hyperbolic_model} and \ref{ssec:local_clustering} of the introduction, we give the definitions of the hyperbolic random graph and of the local clustering coefficient and function as used in this paper. After this we present our main results in Section \ref{ssec:main_results} and discuss several insights that can be obtained from them and in Section \ref{ssec:simulations} show simulations which confirm these results. Section \ref{sec:proof_outline} contains a detailed high level outline of the proofs of our main results, split up into different propositions. The proofs of these propositions can be found in the last four sections. Section~\ref{sec:concentration_argument} contains a crucial technical result that will be used in the other sections. Finally, the Appendix contains a few general results for the hyperbolic random graph and a small section on Meijer's G-function.

@@

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
All our results for clustering in the hyperbolic random graph are valid for any sequence $k_n$ such that $k_n = \bigO{n^{\frac{1}{2\alpha + 1}}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the optimal scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_local_clustering_general} we have that $c_{\H,n}(k_n) = 0$ if $N_{\H,n}(k_n) = 0$. Hence, it follows by Markov's inequality that for any positive function $f$
\begin{align*}
	\Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	&\ge \Prob{N_{\H,n}(k_n) = 0} \ge 1 - \Exp{N_{\H,n}(k_n)}.
\end{align*}
We shall later establish (see Lemma~\ref{lem:N_k_HP_P}) that $\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}$. Therefore if $k_n$ is such that $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$ we have
\[
	\lim_{n \to \infty} n k_n^{-(2\alpha + 1)} 
	= \lim_{n \to \infty} \left(n^{-\frac{1}{2\alpha + 1}} k_n\right)^{-(2\alpha + 1)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	\ge \lim_{n \to \infty} 1 - \Exp{N_{\H,n}(k_n)} = \lim_{n \to \infty} 1 - \bigT{n k_n^{-(2\alpha + 1)}} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:clustering_coefficient_hyperbolic} to hold as soon as $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$, and hence we need $k_n = \bigO{n^{\frac{1}{2\alpha + 1}}}$.

### 2018-10-31 proof_overview.tex
fully merged compations of Nikolaos plus some small edits on section 2

\section{Overview of the proof strategy for Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}}\label{sec:proof_outline}

We first give a detailed overview of the structure for the proof of Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}. The proof itself can be found in Section~\ref{ssec:proof_main_results}, where we also prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}. Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic} based on the different propositions described below, plus the sections in which theses propositions are proved.

>>>>

\section{Overview of the proof strategy}\label{sec:proof_outline}

The key result in this paper is Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}. In particular, it is the key ingredient in the proofs of Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:asymptotics_average_clustering_P}. Therefore, in this section we first give a detailed overview of the structure for the proof of Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}. The proof itself, based on intermediate propositions, can be found in Section~\ref{ssec:proof_main_results}. Here we also prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The proves of Theorem~\ref{thm:exact_expression_c_infty} and Theorem~\ref{thm:exact_expression_local_clustering} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P}

Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic} based on the different propositions described below, plus the sections in which theses propositions are proved.
